---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
author: "Jennyfer Combariza"
date: "December 5, 2016"
---


## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")

```

## What is mean total number of steps taken per day?

```{r}

library(plyr)
library(dplyr)
library(ggplot2)

total.steps <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)

#Mean
mean(total.steps)

#Median
median(total.steps)
```

## What is the average daily activity pattern?

We make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) and we plot the result.


```{r}
five_minutes_average <- aggregate(steps~interval, data=data, FUN=mean, na.rm=TRUE)
plot(x = five_minutes_average$interval, y = five_minutes_average$steps, type = "l") 
png("average.png", width=750)
plot(x = five_minutes_average$interval, y = five_minutes_average$steps, type = "l") 
dev.off()

max_steps <- max(five_minutes_average$steps)
for (i in 1:288) 
{
  if (five_minutes_average$steps[i] == max_steps)
    five_minute_interval_at_max_steps <- five_minutes_average$interval[i]
}
five_minute_interval_at_max_steps 

```

## Imputing missing values

```{r}
sum(!complete.cases(data))


fillfunc <- function(step, interval) {
  ifelse(is.na(step), averages[averages$interval == interval, ]$steps, step)
}

data_fill <- data
tot_steps_day <- aggregate(steps ~ date, data=data_fill, FUN=sum,
                           na.rm=TRUE, na.action=NULL)

granularity = diff(range(tot_steps_day$steps)) / 25
ggplot(data=tot_steps_day, aes(x=tot_steps_day$steps))              +
  geom_histogram(binwidth=granularity, col="black", fill="white") +
  labs(x="Total Steps (with imputed missing data)", y="Count")    +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8, face="italic"))

```

## Are there differences in activity patterns between weekdays and weekends?

We start by creating a new factor variable in the dataset with two levels -- weekday and weekend indicating whether a given date is a weekday or weekend day.

```{r}

switch(Sys.info()[[ 'sysname' ]],
    Windows = { lctime <- "English" }, { lctime <- "C" })
Sys.setlocale("LC_TIME", lctime)

data_fill$date <- as.Date(data_fill$date, "%Y-%m-%d")
daytype <- function(day) {
    ifelse(weekdays(day, abbreviate=FALSE) %in% c("Saturday", "Sunday"),
           "weekend", "weekday")
}
data_fill$daytype <- mapply(daytype, data_fill$date)

```

And we end this document by plotting the two resulting datasets:

```{r}
averages <- aggregate(steps ~ interval + daytype, data=data_fill, FUN=mean)
ggplot(data=averages, aes(x=interval, y=steps))   +
       geom_line()                                +
       facet_grid(daytype ~ .)                    +
       ggtitle("Average daily activity patterns") +
       xlab("5-Minute Interval")                  +
       ylab("Number of Steps")                    +
       theme(axis.text=element_text(size=8),
             axis.title=element_text(size=8, face="italic"))
```

Answer 

Yes, there are  differences in activity patterns between weekdays and weekends.
